---
title: "608: hw1"
author: "Jie Zou"
date: "`r Sys.Date()`"
output:
  html_document: default
  html_notebook: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(gghighlight)
```


  not sure which year is the data being collected, I did a little research about the top rank company. If it is around 2013, according to Wikipedia, its growth rate makes sense.
```{r}
# from the summary, Growth_Rate seems to have outliers

inc[1:5,]
```


  less than 1% of data present with NA, for the further analysis, I will decide to drop them.
```{r}
# from summary, Employees have missing value
data <- inc %>% 
  filter(!is.na(Employees))

```


  Total 25 different industries in which IT service take up the most, which is the most popular one
```{r}
# Among all state, which industry is the most popular?

data %>% 
  select(Industry) %>% 
  group_by(Industry) %>% 
  count(sort = T)
  

```
  based on extracted data, Business Products & Services has the highest revenue across states
```{r}
# which industry has the most annual revenue?

data %>% 
  select(Industry, Revenue) %>% 
  group_by(Industry) %>% 
  summarize(total_revenue = sum(Revenue)) %>% 
  arrange(desc(total_revenue))
  
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
# subset of data with only state and number of companies
states <- data %>% 
  group_by(State) %>% 
  select(Name, State) %>% 
  count(Name) %>% 
  summarize(num_comp = sum(n)) %>% 
  arrange(desc(num_comp))

# first few row of data
head(states)

# make barplot and highlight the top three
states %>% 
  ggplot(aes(x = State, y = num_comp, fill = State)) + 
  geom_bar(stat = 'identity') +
  geom_text(aes(label = num_comp), vjust = 0) + 
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.3, hjust = 0.3)) +
  gghighlight(State == 'CA' | State == 'TX' | State == 'NY') +
  labs(x = 'States',
       y = 'number of companies',
       title = 'distribution of company amount in states')
  
```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}
# get the complete cases in NY state
ny <- data %>% 
  filter(State == 'NY') %>% 
  filter(complete.cases(.))

# make box plot to show median and range. scale to reduce extreme outliers
ny %>% 
  select(Industry, Employees) %>% 
  ggplot(aes(x = Industry, y = Employees)) +
  geom_boxplot() +
  scale_y_log10() + 
  coord_flip() + 
  theme(panel.background = element_blank()) +
  labs(x = 'Industry (by median)',
       y = 'log(Employees)',
       title = 'NY employees by different industry')

```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}
# create data with industry and revenue per employees
inv <- ny %>% 
  select(Industry, Employees, Revenue) %>% 
  group_by(Industry) %>% 
  summarize(rev_per_emp = sum(Revenue) / sum(Employees)) %>% 
  arrange(desc(rev_per_emp))

# make barplot and highlight the top one.
inv %>% 
  ggplot(aes(x = Industry, y = rev_per_emp, fill = Industry)) +
  geom_bar(stat = 'identity') +
  theme(panel.background = element_blank()) +
  coord_flip() +
  gghighlight(rev_per_emp == max(inv$rev_per_emp)) +
  labs(x = 'Industry in NY',
       y = 'revenue per employees',
       title = 'NY state industries generate revernue per employees')
```


