---
title: "a_2"
author: "Jie Zou"
date: "2/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# connect postgres and make queries

1. we need to install packages to work on connection, note that RPostgreSQL package is not working in my current version of R, therefore, RPostgres will be the alternative way.

```{r}
install.packages("RPostgre")
```

2. use two libraries under this package 

```{r}
library(DBI)
library(RPostgres)
```

3. connect with the database created in Postgres server and list out all the tables under this DB

```{r}
con<-dbConnect(RPostgres::Postgres(), dbname="top_10_movies_survey")
dbListTables(con)
```

4. make a query to present the table

```{r}
tb = dbGetQuery(con, "select * from moviesurvey")
tb
dbGetQuery(con, "select * from moviesurvey where the_queens_gambit = 5")

```

5. if you would like to work on database locally, you can create data frame based on the table

```{r}
movie_survey <- as.data.frame(tb)
View(movie_survey)
```









