---
title: "p3: what is the most valuable skill in data science"
subtitle: "Analysis and visualization"
author: "Jie Zou"
date: "`r Sys.Date()`"
output: html_document
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(DT)
library(dplyr)
library(plotly)
```

```{r}
df <- read.csv("https://raw.githubusercontent.com/Sugarcane-svg/R/main/R607/Projects/p3/Project3_Progress_210325.csv")

df <- df %>%
  rename(
    position = Position,
    location = Location,
    company = ad_poster,
    company_type = ad_poster_type,
    industry = job_Knowledge_Domain
    )

library(stringr)
df$position <- str_trim(df$position)
```

*This job name also represents data science, so when so many names are at work, which ones appear most often?*
```{r}
positions <- df %>% 
  count(position)

ggplotly(ggplot(positions %>% filter(n > 3), aes(x = position, y = n)) +
  geom_bar(stat = "identity") + 
  coord_flip() +
  labs(x = "number of position",
       y = "name of position",
       title = "Names of position that more than three companies posted "))
```
*What are the difference of annual salary between these companies?*

the subset of data set filters out the the salary paid by day, so we do not need to worry if there are going to be some extreme outliers. I subtract the min_salary from max_salary and square the residual just to make the large value larger in the way to easily see the difference. 
```{r}
# subset of data: [position], [company], [diff_annual_salary], [company_type]
avg_annum <- df %>%
  filter(Salary.Period == "per annum" & !is.na(Salary.Period)) %>%
  mutate(diff_annual_salary = (Max.Salary.Lbs - Min.Salary.Lbs)^2)%>%
  select(position, diff_annual_salary, company, company_type) %>%
  arrange(desc(diff_annual_salary))

view(avg_annum)
# number of companies offer the same average annual salary
count_company <- avg_annum %>% 
  count(diff_annual_salary) %>% 
  arrange(desc(n))

ggplotly(
  ggplot(count_company, aes(x = avg_annual_salary, y = n)) + 
  geom_bar(stat = "identity") +
  labs(y = "number of company",
       title = "annual salary difference within the companies")) 
```

*the larger average, the larger span. which companies offer the largest and the smallest annual salary span?*

according to the calculation, both Harnham and McGregor Boyall provide the largest salary span to the position of Data Scientist and software developer. It seems reasonable. However, Hudson Shribman offers the least salary span in chemist/Analytical chemist/Materials scientist.
```{r}
# min and max of average annual salary
a<-avg_annum %>% filter(diff_annual_salary == max(diff_annual_salary) | diff_annual_salary == min(diff_annual_salary))
datatable(a)
# min and max span of these companies
b<- df %>% filter(((Max.Salary.Lbs-Min.Salary.Lbs) == max(Max.Salary.Lbs-Min.Salary.Lbs) | (Max.Salary.Lbs-Min.Salary.Lbs) == min(Max.Salary.Lbs-Min.Salary.Lbs)) & Salary.Period == "per annum") %>% select(-c(Posted, location, Contract,industry, X, Salary.Period))
datatable(b)
```

*What are the major types and corresponding ratio of these companies?*

the majority type of job posting companies is recruitment consultancy
```{r}
library(scales)
c<-df %>% 
  count(company_type) %>%
  mutate(per = n/sum(n), 
         label = str_c(percent(per), "(", n, ")")) %>%
  arrange(desc(n))

ggplot(data = c) +
  geom_bar(aes(x = "", y = per, fill = company_type), stat = "identity", width = 1) + 
  coord_polar("y", start = 0) +
  geom_text(aes(x = 1.7, y = cumsum(per)-per/2, label = label)) +
  theme_void()
```
*Among recruitment consultancy, which company or which companies have the most influence?*
The company is considered influential, so its business scope should be broad. In other words, a company like this has more cooperative work than other companies, which means it will post more jobs. As we can see from the plot, the number of jobs posted by Harnham far exceeds that of other companies
```{r}
# business with wider scope
d1 <- df %>% 
  filter(company_type == "Recruitment consultancy") %>%
  count(company) %>% 
  arrange(desc(n))

ggplot(d1 %>% filter(n > 2), aes(x = company, y = n)) + 
  geom_bar(stat = "identity")+
  coord_flip() +
  labs(y = "number of job posts",
       title = "recruitment consultancy with at least 2 job posts")
```
*How is the spread of locations?*

we've select the companies with at least 4 job posts in one location. Ro discover the relationship between companies and location, the scatter plot is shown below, where we can see that most of companies have job location in London, some companies provide more than one locations such as Harnham and Blue Pelican.
```{r}
# location of these job distributed by companies with at least 4 job posts
d2 <- df %>% 
  group_by(company)%>%
  count(location) %>%
  filter(n > 3)
  
ggplot(d2, aes(x = location, y = company)) +
  geom_point(aes(color = location)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

