---
title: 'a_4: Tidying and Transforming Data'
author: "Jie Zou"
date: "3/4/2021"
output:
  html_document: default
  openintro::lab_report: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(knitr)
```

## Create a .CSV file

The data from provided image is pretty small, therefore I manually enter these data into Excel with exactly the same format showed in the image, and save it as .CSV file and upload it into git first.


## Read .CSV file

```{r}
df<-read.csv("https://raw.githubusercontent.com/Sugarcane-svg/R/main/R607/Assignments/a4/a_4.csv")
kable(df)

```



## tidy data

To tidy data, we are going to make up the missing column names, remove na or unnecessary 'white line', and make up some data(if necessary)

#### 1. modify column names
```{r}
colnames(df) <- c("airline", "status","LA", "PHX", "SAN", "SFO", "SEA")
head(df)
```

#### 2. remove line 3 

```{r}
df <- df[-c(3),]
head(df)
```

#### 3. fill data in blank space

```{r}
df[2,1] <-"ALASKA"
df[4,1] <-"AMWEST"
head(df)
```

#### 4(optional). split data into two for the simplicity of following steps
```{r}
alaska.df<-data.frame(airline = character(), dest = character(), on_time = numeric(), delayed = numeric())

amwest.df<-data.frame(airline = character(), dest = character(), on_time = numeric(), delayed = numeric())

for( i in 1:5){
  alaska.df[i,] <- c("ALASKA", 
                     colnames(df)[i+2],
                     (df %>%
                       filter(airline == "ALASKA" & status == "on time")%>%
                       select(-c(airline, status)))[1,i],
                     
                     (df %>%
                       filter(airline == "ALASKA" & status == "delayed")%>%
                       select(-c(airline, status)))[1,i]
                     )
}

alaska.df
for( i in 1:5){
  amwest.df[i,] <- c("AMWEST", 
                     colnames(df)[i+2],
                     (df %>%
                       filter(airline == "AMWEST" & status == "on time")%>%
                       select(-c(airline, status)))[1,i],
                     
                     (df %>%
                       filter(airline == "AMWEST" & status == "delayed")%>%
                       select(-c(airline, status)))[1,i]
                     )
}
amwest.df

```

## Analysis

### compare the arrival delays for two airlines

#### 1. lets see the mean of both on time and delayed values from these two airlines
```{r}
df$mean <- rowMeans(df[, 3:7])
df
```
as we can see from the mean, it seems like Alaska has better on time and less delayed. However, when we observe the data, we can also see that in some specific destinations, there are some on time flights unexpected longer than others such as flying from Alaska to Seattle and flying from Amwest to Phoenix. The number of both on time are much higher compared to other destinations.


#### 2. check the ratio of arrival delays for both airline

The total number of flights for Alaska is 3775 and Amwest is 7225
```{r}
df$total <- rowSums(df[, 3:7])
df %>% select(airline, total) %>%
  group_by(airline) %>%
  summarise(total_number_of_flights = sum(total))
```

the probability of delay for alaska and amwest flights are 13.27% and 10.89% respectively. compare the ratio, we can see that the delay from alaska is worse than the delay from amwest.
```{r}

# alaska delay ratio
ak_delay <-df %>% 
  filter(airline == "ALASKA" & status == "delayed") %>%
  select(total)

ak_delay/3775


# amwest delay ratio

am_delay<- df %>% 
  filter(airline == "AMWEST" & status == "delayed") %>%
  select(total)

am_delay/7225

```


#### 3. see the graph of individual airline delay status of listed city

from the graph, it can only tell/confirm that there are more flights delayed from Alaska to Seattle, and there are more flights delayed from Amwest to Phoenix. The flight delayed from Alaska to San Diego is less than the flights from Amwest. 

```{r}
ak_delay <- alaska.df %>% select(dest, delayed) %>% arrange(dest)

ggplot(ak_delay, aes(x=dest, y = as.numeric(delayed))) + geom_bar(stat = "identity") + labs(x = "destination", y = "number of delayed flights", title = "Alaska flights delay status to listed city")

am_delay <- amwest.df %>% select(dest, delayed)%>% arrange(dest)

ggplot(am_delay, aes(x=dest, y = as.numeric(delayed))) + geom_bar(stat = "identity") + labs(x = "destination", y = "number of delayed flights", title = "Amwest flights delay status to listed city")
```


#### 4. see both airline delay status in one graph

according to this graph, it shows clear comparison between these two airlines to selected five cities. As we can see that flights from Alaska to LA, PHX, SAN, SFO have less number of delayed. And flying from Amwest to Seattle appears less delayed.
```{r}
df2 <- rbind(alaska.df, amwest.df)
df2

ggplot(df2, aes(x=dest, y = as.numeric(delayed), fill = airline)) +
  geom_bar(stat = "identity")+
  labs(x = "destination", y = "number of delayed flights", title = "delay status to listed city")
  
```

## Conclusion

If a customer is deciding flying from these two airline to San Diego/Phoenix, I will suggest to take Alaska, and if a customer try to find a airline to Seattle, I will suggest to choose Amwest.